---
title: "Reproducible Research from JHU, Assignment1"
author: "Steve Laurie"
date: "1 February 2016"
output: html_document
---
Load dplyr and lattice which we will use for "filter" and plotting.
```{r}
library(dplyr)
library(lattice)
```

Load the data and have a first look.

```{r}
rawdata<-read.csv("activity.csv")
dim(rawdata)
head(rawdata)
```

How many days do we have data for?

```{r}
length(unique(rawdata$date))
```

Remove rows that have "NA"

```{r}
Good <- complete.cases(rawdata)
cleandata <- (rawdata[Good,])
dim(cleandata)
length(unique(cleandata$date))
```

We can see that we have removed 8 entire days for which there was no data, leaving us with 53 for further analysis.


### **What is the mean number of total steps taken per day?**

Calculate total number of steps per day
```{r}
splitByDay<-split(cleandata, cleandata$date)
sapply(splitByDay, function(x){sum(x[, "steps" ])})

```

Visual inspection of the table shows that that the greatest number of steps per day was just over 21k, therefore going to generate a histogram with 22 columns, each representing a 1000 steps.
```{r}
hist(sapply(splitByDay, function(x){sum(x[, "steps" ])}),  breaks=21, main="Histogram of number of steps taken per day over a 61-day period", xlab="Steps taken per day", ylab="Number of days")
```

Calculate mean and median number of steps per day
```{r}
summary(sapply(splitByDay, function(x){sum(x[, "steps" ])}))
```

We can see from the summary table above that the **median number of steps per day is 10,400, while the mean is 9,354**. The low mean is due to the large number of days (10), where the total number of steps was less than 1,000 (typically 0).

## Make a time-series plot

Here we are asked to make a time series plot for each moment of the day (across all the days) to see when most steps are taken. So, we can divide this into two steps:

* Get the sum for each of the 288 5-min segements in a day
* Generate and plot the time series

Reusing from above
```{r}
splitByInterval<-split(cleandata, cleandata$interval)
SumByInterval<-sapply(splitByInterval, function(x){sum(x[, "steps" ])})
```

Generate the plot

```{r}
IntervalTimeSeries<-ts(SumByInterval)
plot.ts(IntervalTimeSeries, xy.labels=cleandata$interval)
```

It is clear from the time-series plot that there is a peak in the early morning, around about interval 105. 

```{r}
which.max(IntervalTimeSeries)
```

This shows us that it is exactly interval 104, which corresponds to 835-084a.m.

## Imputing Missing Values

### **Calculate and report the total no. of missing values in the dataset**

We already have this from above when we looked for the complete.cases
```{r}
table(Good)
```


This indicates that **there are 2304 rows with missing data** i.e. "NAs"

### Devise strategy for filling in all of the missing values - does not need to be sophisticated

I am going to use just the median value for each interval

```{r}
MedianByInterval<-sapply(splitByInterval, function(x){median(x[, "steps" ])})
```

I found a video that illustrates how to do this:
https://www.youtube.com/watch?v=w1t1a_doQoo

Identify the rows which have an NA in the steps column, and replace with the median value for that inteval, in a new dataframe named Impute.

```{r eval=FALSE}
which(is.na(rawdata$steps))
```
```{r}
MedianRep<-rep(MedianByInterval,61)
Imputed <-cbind.data.frame(rawdata, Median = MedianRep)
Imputed$steps[is.na(Imputed$steps)] <- Imputed$Median[is.na(Imputed$steps)]
splitByDay2<-split(Imputed, Imputed$date)
```

Make a hist of the new data:

```{r}
hist(sapply(splitByDay2, function(x){sum(x[, "steps" ])}),  breaks=21, main="Histogram of number of steps taken per day following imputation, over a 61-day period", xlab="Steps taken per day", ylab="Number of days")
```

Check what the mean/median is per day using this new dataset. Main difference appears to be that a lot of the values formerly in bin1 have moved into bin2 - i.e. 1,000-2,000 steps per day.



```{r}
summary(sapply(splitByDay2, function(x){sum(x[, "steps" ])}))
```

Naturally the median hasn't changed, since this was what we used for imputing, but the mean has increased from 9354 to 9504 steps per day. 

```{r}
sapply(splitByDay2, function(x){sum(x[, "steps" ])})
```

# What about weekends? #

Update dataframe and split in two

```{r}
Imputed2 <-cbind.data.frame(Imputed, Day = weekdays(as.Date(Imputed$date, abbr=T)))
Imputed2$Weekday <- ifelse(Imputed2$Day == 'lunes'|Imputed2$Day =='martes'|Imputed2$Day =='miércoles'|Imputed2$Day =='jueves'|Imputed2$Day =='viernes',"weekday","weekend")
Weekdays.df<-filter(Imputed2, Weekday=='weekday')
Weekends.df<-filter(Imputed2, Weekday!='weekday')
```
Get summary for each  
```{r}
summary(sapply(split(Weekends.df, Weekends.df$date), function(x){sum(x[, "steps" ])}))
summary(sapply(split(Weekdays.df, Weekdays.df$date), function(x){sum(x[, "steps" ])}))
```

```{r}
WeekdaysByInterval<-split(Weekdays.df, Weekdays.df$interval)
SumByForWeekdays<-sapply(WeekdaysByInterval, function(x){sum(x[, "steps" ])})
WeekdaysInterval.ts<-ts(SumByForWeekdays)

WeekendsByInterval<-split(Weekends.df, Weekends.df$interval)
SumByForWeekends<-sapply(WeekendsByInterval, function(x){sum(x[, "steps" ])})
WeekendsInterval.ts<-ts(SumByForWeekends)

plot.ts(WeekdaysInterval.ts)
plot.ts(WeekendsInterval.ts)
```
  
Make the plots - maybe check http://www.fromthebottomoftheheap.net/2013/10/23/time-series-plots-with-lattice-and-ggplot/
  
```{r}
Imputed2 <- transform(Imputed2, Weekday = factor(Weekday))
```
  
```{r}
class(Imputed2$Weekday)
```


```{r}
xyplot(
  ts(sapply(split(filter(Imputed2, Weekday!='weekday'), Imputed2$interval)
, function(x){sum(x[, "steps" ])})) ~
ts(sapply(split(filter(Imputed2, Weekday=='weekday'), Imputed2$interval)
, function(x){sum(x[, "steps" ])})) | Imputed2$Weekday, layout = c(2, 1)
)
```
  
  That's not what I want  
  
```{r}
xyplot(
  sapply(split(Imputed2, Imputed2$date), function(x){sum(x[, "steps" ])}) ~ Imputed2$interval| Imputed2$Weekday, layout =c(1,2)
)
```
  
   
  Neither is this. I have spent a lot of time on this, and I am obviously doing something silly, so I'll end here.  

**Thanks for your time marking **
